---
pagetitle: Distributions
output: 
  revealjs::revealjs_presentation:
    incremental: false
    theme: solarized
    self_contained: false
    # reveal_plugins: ["menu","notes","chalkboard"]
    reveal_plugins: ["menu"]
    highlight: pygments
    center: true
    transition: none
    background_transition: none 
    reveal_options:
      # chalkboard:
      #   theme: whiteboard
      #   toggleNotesButton: true
      #   toggleChalkboardButton: true
      menu:
        numbers: true
      slideNumber: true
      previewLinks: false
    fig_caption: true
    css: mystyle.css
    
--- 

<section>

<h1>Distributions</h1>

Based on Stock and Watson, ch. 2

<br>

<h2>[Jesper Bagger](mailto:jesper.bagger@rhul.ac.uk)</h2>

<h3>EC2203 | Royal Holloway | 2020/21</h3>

</section>

# Distributions

## Probability

- Suppose $X$ is a random variable; the probability that $X=a$ is the **proportion of times $X$ take on the value $a$ in many, many independent realizations** of $X$

- For example, consider the random variable $X$ constructed from a coin flip with potential outcomes $\{H,T\}$

  $$X = \left\{
  \begin{array}{ll}
  1 & \text{if } H, \\
  0 & \text{if } T;
  \end{array}
  \right.$$

  
  then, $P(X = 1)$ is the proportion of times that $X=1$ occurred the in many, many coin-flip repetitions


## Distributions

- The **distribution** of a random variable specifies the probability of realizing *each* value of the random variable

- Distinguish b/w **discrete** and **continuous** random variables

## The probability mass function 

- The **probability mass function (PMF)** of a **discrete** random variable $X$ is a function $f_X$ such that

  <div class="box">
  $$f_X(a) = P(X = a); \quad  \sum_a f_X(a) = 1$$
  </div>
  
- The PMF is a list of values taken by the random variable $X$ together with a list of associated probabilities 

## The cumulative distribution function

- The **cumulative distribution function (CDF)** of a random variable $X$ is a function $F_X$ such that

  <div class="box">
  $$F_X(a) = P(X \leq a)$$
  </div>
  

<!-- ## The CDF and PMF of discrete random variables -->

<!-- - When $X$ is a discrete random variable, $F_X$ is the cumulative sum of $f_X$ -->

<!--   <div class="box"> -->
<!--   $$F_X(x) = \sum_{z \leq x} f_X(z)$$ -->
<!--   </div> -->
  
## The CDF and PDF of a continuous random variable

- The definition of the cumulative distribution function extends to **continuous** random variables

- The PMF is unfit for continuous random variables; instead, introduce the **probability density function (PDF)**

  <div class="box">
  $$f_X(a) = \frac{d}{da} F_X(a)$$
  </div>

- The area under the PDF $f_X$ between two points is equal to the probability that $X$ takes a value between those points

  


## The Normal PDF and CDF

```{r echo=FALSE, error=FALSE, warning=FALSE}
par(mfrow=c(2,1), mar = c(2,4.1,1,4.1))
# draw a plot of the N(0,1) PDF
x <- seq(-3.5,3.5,length=200)
pdf <- dnorm(x)
plot(x,pdf,
      type = "l",
      cex.axis = 1.25, 
      cex.lab = 1.25, 
      xlim = c(-3.5, 3.5),
      ylab = "Density",
      xlab = " ",
      col = "blue",
      lwd = 3,
      panel.first = grid(nx = NULL, 
                        ny = NULL, 
                        col = "lightgray", 
                        lty = "dotted", 
                        lwd = 2)) 
box(which = "plot", 
    lty = "solid", 
    lwd = 3)
# draw a plot of the N(0,1) CDF
x <- seq(-3.5,3.5,length=200)
cdf <- pnorm(x)
plot(x,cdf,
      type = "l",
      cex.axis = 1.25, 
      cex.lab = 1.25, 
      xlim = c(-3.5, 3.5),
      ylab = "Probability",
      xlab = " ",
      col = "blue",
      lwd = 3,
      panel.first = grid(nx = NULL, 
                        ny = NULL, 
                        col = "lightgray", 
                        lty = "dotted", 
                        lwd = 2)) 
box(which = "plot", 
    lty = "solid", 
    lwd = 3)

```


## The Normal PDF and CDF

```{r echo=FALSE, error=FALSE, warning=FALSE}
par(mfrow=c(2,1), mar = c(2,4.1,1,4.1))
# draw a plot of the N(0,1) PDF

x <- seq(-3.5,3.5,length=200)
pdf <- dnorm(x)
plot(x,pdf,
      type = "l",
      cex.axis = 1.25, 
      cex.lab = 1.25, 
      xlim = c(-3.5, 3.5),
      ylab = "Density",
      xlab = " ",
      col = "blue",
      lwd = 3,
      panel.first = grid(nx = NULL, 
                        ny = NULL, 
                        col = "lightgray", 
                        lty = "dotted", 
                        lwd = 2)) 
box(which = "plot", 
    lty = "solid", 
    lwd = 3)

# The shaded area
xshade <- seq(-1.96,1.96,length=200)
pdfshade <- dnorm(xshade)
polygon(c(-1.96,xshade,1.96),c(0,pdfshade,0), col="lightgray", lwd = 0.1)
lines(x,pdf,
      lwd = 3,
      col="blue")

text(x = 0,y = 0.15,"Area: 0.95")

# draw a plot of the N(0,1) CDF
x <- seq(-3.5,3.5,length=200)
cdf <- pnorm(x)
plot(x,cdf,
      type = "l",
      cex.axis = 1.25, 
      cex.lab = 1.25, 
      xlim = c(-3.5, 3.5),
      ylab = "Probability",
      xlab = " ",
      col = "blue",
      lwd = 3,
      panel.first = grid(nx = NULL, 
                        ny = NULL, 
                        col = "lightgray", 
                        lty = "dotted", 
                        lwd = 2)) 
box(which = "plot",
    lty = "solid",
    lwd = 3)

# segments(1.96, -1, 1.96, pnorm(1.96), col = "black", lty = 2, lwd = 3)
# segments(-4, pnorm(1.96), 1.96, pnorm(1.96), col = "black", lty = 2, lwd = 3)
# mtext('0.975 ', side=2, at=c(-4.5,pnorm(1.96)),las = 2)
# 
# segments(-1.96, -1, -1.96, pnorm(-1.96), col = "black", lty = 2, lwd = 3)
# segments(-4, pnorm(-1.96), -1.96, pnorm(-1.96), col = "black", lty = 2, lwd = 3)
# mtext('0.025 ', side=2, at=c(-4.5,pnorm(-1.96)),las = 2)
```


# Summary

## Summary

- $P(X = a)$, is the proportion of times $X$ takes the value $a$ in many independent realizations of $X$

- The distribution of a random variable specifies the probability for each value taken by the random variable

- The probabilities with which a discrete random variable takes on different values are given by the PMF

- The probabilities with which a continuous random variable takes on different values are given by the PDF

- The CDF gives the probability that a random variable takes a value less than or equal to a specific value

